---
title: "Scurrying Secrets: Unraveling the Tapestry of Squirrel Activity"
---

we could also name this "Acorn Adventures: Unraveling the Tapestry of Squirrel Activity"

tables and whatever about "what squirrels do"

```{r include = FALSE}
library(tidyverse)
library(plotly)
library(viridis)
```

```{r import_data, message = FALSE}
squirrel_census = 
  read_csv("data/clean_squirrel_2018.csv")
```

Squirrels are dynamic creatures with a secret language. In a bustling environment like New York City, these animals use unique signals to communicate with each other, other animals, and us. 

First, let's take a look at squirrel sounds. 

```{r sound_distribution}
sounds_summary = data.frame(
  Sound = c("Kuks", "Quaas", "Moans"),
  Count = c(sum(squirrel_census[["kuks"]], na.rm = TRUE), 
            sum(squirrel_census[["quaas"]], na.rm = TRUE), 
            sum(squirrel_census[["moans"]], na.rm = TRUE))
)

# Squirrel sounds distribution 
ggplot(sounds_summary, aes(x = Sound, y = Count, fill = Sound)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal() +
  labs(title = "Distribution of Squirrel Sounds", x = "Sound Type", y = "Count")
```

Of the three identified squirrel sounds, NYC squirrels most commonly "kuk" is the most common sound, and "moans" are rarely heard. To better understand the meaning of these sounds, let's take a look at the type of activity or interactions the squirrels who made these sounds are engaged in. 









