---
title: "Scurrying Secrets: Unraveling the Tapestry of Squirrel Activity"
---

we could also name this "Acorn Adventures: Unraveling the Tapestry of Squirrel Activity"

tables and whatever about "what squirrels do"

```{r include = FALSE}
library(tidyverse)
library(plotly)
library(viridis)
```

```{r import_data, message = FALSE}
squirrel_census = 
  read_csv("data/clean_squirrel_2018.csv")
```

Squirrels are dynamic creatures with a secret language. In a bustling environment like New York City, these animals use unique signals to communicate with each other, other animals, and us. 

First, let's take a look at squirrel sounds. 

```{r sound_distribution}
sounds_summary = data.frame(
  Sound = c("Kuks", "Quaas", "Moans"),
  Count = c(sum(squirrel_census[["kuks"]], na.rm = TRUE), 
            sum(squirrel_census[["quaas"]], na.rm = TRUE), 
            sum(squirrel_census[["moans"]], na.rm = TRUE))
)
```

```{r squirrel_sounds}
plot_ly(data = sounds_summary, x = ~Sound, y = ~Count, color = ~Sound, type = 'bar', colors = viridis::viridis(3)) |>
  layout(title = 'Distribution of Squirrel Sounds',
         xaxis = list(title = 'Sound Type'),
         yaxis = list(title = 'Count'))

```

Of the three identified squirrel sounds, NYC squirrels most commonly "kuk" is the most common sound, and "moans" are rarely heard. To better understand the meaning of these sounds, let's take a look at the type of activity or interactions the squirrels who made these sounds are engaged in. 

```{r}
# Summarizing sounds by interaction types
interaction_sounds <- squirrel_census |>
  group_by(approaches, indifferent, runs_from) |>
  summarise(Kuks = sum(kuks, na.rm = TRUE),
            Quaas = sum(quaas, na.rm = TRUE),
            Moans = sum(moans, na.rm = TRUE)) |>
  pivot_longer(cols = c(Kuks, Quaas, Moans), names_to = "Sound", values_to = "Count")

# Creating a new column for interaction type
interaction_sounds[["interaction_type"]] <- case_when(
  interaction_sounds[["approaches"]] == TRUE ~ "Approaches",
  interaction_sounds[["indifferent"]] == TRUE ~ "Indifferent",
  interaction_sounds[["runs_from"]] == TRUE ~ "Runs From",
  TRUE ~ "Other"
)

# Creating sounds by human interaction plot
plot_ly(interaction_sounds, x = ~Sound, y = ~Count, color = ~interaction_type, type = 'bar', colors = RColorBrewer::brewer.pal(3, "Set2")) |>
  layout(title = 'Distribution of Squirrel Sounds by Human Interaction',
         xaxis = list(title = 'Sound Type'),
         yaxis = list(title = 'Count'))
```

As the plot above shows, squirrels who `kuk`are frequently `indifferent,` `run from,` or exhibit `other` behavior when interacting with humans. Squirrels who `quaa` also are `indifferent` or exhibit `other` behavior when approached by a human, and `run from` humans less frequently. There is only one instance of a `moan`, and this squirrel `ran from` a human. While this is an interesting data point, there is not enough information to associate a `moan` with fear or distress. Without more data and a better understanding of what `other` behavior is, we cannot accurately determine what these squirrel sounds mean. 


