---
title: "Squirreling the Stats: A Demographic Tale"
output: 
  html_document:
    code_folding: hide
    always_allow_html: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(plotly)
library(sf)
```

#### In this section, we will be looking at squirrel demographics, such as fur color and age. 

First, we will import the cleaned squirrel data and a shapefile of Central Park. 

```{r}
squirrel_census = 
  read_csv("data/clean_squirrel_2018.csv")

central_park_map = st_read(here::here('central_park', 'CentralPark.shp'), quiet = TRUE)
```

There are a total of `r nrow(squirrel_census)` squirrel observations in the data set. 


#### Let's first explore squirrel ages. 

```{r, fig.width= 6, fig.align='center'}
squirrel_census |> 
  filter(age != "?") |> 
  ggplot() +
  geom_sf(data = central_park_map, color = "grey") +
  geom_point(
    aes(x = x, y = y, color = age),
    size = 0.5, alpha = 0.75) +
    theme_void(base_size = 15) +
  theme(legend.position = 'bottom') +
  guides(color = guide_legend(
    title.position = "top",
    override.aes = list(size = 3))) +
  scale_color_manual(values = c("Adult" = "aquamarine3", "Juvenile" = "coral1")) +
  labs(title = "Squirrel Age",
       color = "Age")
```

```{r}
squirrel_census |> 
  filter(!(age %in% c("N/A", "?"))) |> 
  group_by(age) |> 
  count(age) |> 
  mutate(Age = fct_reorder(age, n)) |> 
  plot_ly (x = ~age, y = ~n, color = ~age,
         type = "bar") |> 
  layout(
    xaxis = list(title = "Age"),   
    yaxis = list(title = "Number of Squirrels") 
  )
```
It appears that the majority of observed squirrels are adults. There are 2503 adult squirrels in the data set and only 319 juvenile squirrels. 

#### Let's also look at primary and highlight fur color.


```{r, fig.width= 6, fig.align='center'}
ggplot() +
  geom_sf(data = central_park_map, color = "grey") +
  geom_point(
    data = squirrel_census,
    aes(x = x, y = y, color = primary_fur_color),
    size = 0.5, alpha = 0.75) +
    theme_void(base_size = 15) +
  theme(legend.position = 'bottom') +
  guides(color = guide_legend(
    title.position = "top",
    override.aes = list(size = 3))) +
  labs(color = 'Primary Fur Color',
       title = 'Squirrel Fur Color Distribution') +
  scale_color_manual(values = c("Black" = "black", "Gray" = "azure4", "Cinnamon" = "darkorange3")) 
```


```{r}
squirrel_census |> 
  filter(!(combination_of_primary_and_highlight_color %in% c("N/A", "?"))) |> 
  group_by(primary_fur_color) |> 
  count(primary_fur_color) |> 
  mutate(primary_fur_color = fct_reorder(primary_fur_color, n)) |> 
  plot_ly (x = ~primary_fur_color, y = ~n, color = ~primary_fur_color,
         type = "bar", colors = "viridis") |> 
  layout(
    xaxis = list(title = "Primary Fur Color"),   
    yaxis = list(title = "Number of Squirrels") 
  )
```

```{r}
squirrel_census |> 
  rename(combination_fur_color = combination_of_primary_and_highlight_color) |> 
  count(combination_fur_color) |> 
  mutate(combination_fur_color = fct_reorder(combination_fur_color, n)) |> 
  plot_ly (x = ~combination_fur_color, y = ~n, color = ~combination_fur_color,
         type = "bar", colors = "viridis") |> 
  layout(
    xaxis = list(title = "Fur Color"),   
    yaxis = list(title = "Number of Squirrels") 
  )
```

#### Is there a difference in primary fur color depending on age? 

```{r}
squirrel_census |> 
  filter(!(age %in% c("N/A", "?"))) |>
  select(age, primary_fur_color) |> 
  group_by(age, primary_fur_color) |> 
  summarise(count = n()) |> 
  mutate(percent = count / sum(count) * 100) |> 
  plot_ly (x = ~age, y = ~percent, color = ~primary_fur_color,
         type = "bar", colors = "viridis") |> 
  layout(
    xaxis = list(title = "Age"),   
    yaxis = list(title = "Percent of Squirrels") 
  )
```
