---
title: "squirrel data clean and explore"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)

library(tidyverse)
library(lubridate)
```

*Summary*

The `clean_squirrel_2018` data is a wide data set with one row per
squirrel. These squirrels have nonmissing data for the most critical
variables for analysis (identifier, geodata, fur color, age, activity,
date).

### Data cleaning check list:

*   every squirrel has a unique ID [x]
*   every squirrel has geodata (`hectare`, `x`, `y`) [x]
*   non-missing data for:
    *   primary fur color [x]
    *   age [x]
    *   location description [x]
    *   activity descriptions [x]
    *   date [x]
*   continuous variables as numeric [x]
*   document number of squirrels that fit all criteria [x]

### Get data

Read in 2018 Squirrel Census data from:
<https://data.cityofnewyork.us/Environment/2018-Central-Park-Squirrel-Census-Squirrel-Data/vfnx-vebw>.

Use skimmr to check variable completeness, types, etc.
```{r}
df_dirty_squirrel <- read_csv("2018_Central_Park_Squirrel_Census_-_Squirrel_Data_20231204.csv")

skimr::skim(df_dirty_squirrel)
```



### Clean variable names and remove missing data
```{r}
df_clean_squirrel = df_dirty_squirrel |> 
  janitor::clean_names() |> 
  filter(
    complete.cases(unique_squirrel_id, 
                   x, y, 
                   primary_fur_color, 
                   age, 
                   location
                   )
  ) |> 
  mutate(
    x = as.numeric(x), 
    y= as.numeric(y), 
    day = as.numeric(substr(date, 3, 4)),     #date results in a lot of parsing failures so I'm using a work around
    month = as.numeric(substr(date, 1, 2)),  
    year = as.numeric(substr(date, 5, 11)), 
    date_assembled = paste(month, day, year, sep = "-") 
  ) |> 
  mutate(
    date = lubridate::mdy(date_assembled)
  ) |> 
  filter(!is.na(date))
```

In the cleaner version of the data for which only squirrels with non-missing data have been kept, there are data for 2819 distinct squirrels. The data was collected between 2018-10-06 and 2018-10-20.

### Export clean data
```{r}
save_path = "clean_squirrel_2018.csv"

write_csv(df_clean_squirrel, save_path, na="")
```
