---
title: "squirrel data clean and explore"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)

library(tidyverse)
library(lubridate)
```

*Summary*

The `clean_squirrel_2018` data is a wide data set with one row per
squirrel. These squirrels have nonmissing data for the most critical
variables for analysis (identifier, geodata, fur color, age, activity,
date).

### Data cleaning check list:

*   every squirrel has a unique ID [x]
*   every squirrel has geodata (`hectare`, `x`, `y`) [x]
*   non-missing data for:
    *   primary fur color [x]
    *   age [x]
    *   location description [x]
    *   activity descriptions [x]
    *   date [x]
*   continuous variables as numeric [x]
*   document number of squirrels that fit all criteria [x]

### Get data

Read in 2018 Squirrel Census data from:
<https://data.cityofnewyork.us/Environment/2018-Central-Park-Squirrel-Census-Squirrel-Data/vfnx-vebw>.

Use skimmr to check variable completeness, types, etc.
```{r}
df_dirty_squirrel <- read_csv("2018_Central_Park_Squirrel_Census_-_Squirrel_Data_20231204.csv")

skimr::skim(df_dirty_squirrel)
```



### Clean variable names and remove missing data
```{r}
df_clean_squirrel = df_dirty_squirrel |> 
  janitor::clean_names() |> 
  filter(
    complete.cases(unique_squirrel_id, 
                   x, y, 
                   primary_fur_color, 
                   age, 
                   location
                   )
  ) |> 
  mutate(
    x = as.numeric(x), 
    y= as.numeric(y), 
    day = as.numeric(substr(date, 3, 4)),     #date results in a lot of parsing failures so I'm using a work around
    month = as.numeric(substr(date, 1, 2)),  
    year = as.numeric(substr(date, 5, 11)), 
    date_assembled = paste(month, day, year, sep = "-") 
  ) |> 
  mutate(
    date = lubridate::mdy(date_assembled)
  ) |> 
  filter(!is.na(date))
```

In the cleaner version of the data for which only squirrels with non-missing data have been kept, there are data for 2819 distinct squirrels. The data was collected between 2018-10-06 and 2018-10-20.

### Adding squirrel names 
```{r}
squirrel_names <- c(
  "Sparky Nuttail", "Cocoa Whiskertail", "Hazel Paws", "Sable Furryfluff", "Almond Bushytoes",
  "Pecan Snugglewhisk", "Popcorn Nutbuster", "Chestnut Snickersnack", "Butterscotch Fluffpaws",
  "Maple Tangletoe", "Walnut Giggletail", "Acorn Snickersnout", "Pistachio Wigglesnout",
  "Cashew Sillytail", "Toffee Fluffernut", "Macadamia Wigglewhisk", "Cinnamon Nibbler",
  "Russet Snickerwhisk", "Caramel Fuzzernose", "Mocha Whiskernib", "S'mores Snickersnout",
  "Olive Gigglenose", "Mango Snuggletail", "Taffy Wigglesnout", "Marshmallow Nibbler",
  "Chestnut Scurrypaws", "Butter Pouncewhisk", "Truffle Fluffpounce", "Pumpkin Nibblesnout",
  "Cocoa Wigglepaws", "Peanut Whiskersnack", "Sesame Nuzzlenib", "Blueberry Fluffernibble",
  "Hazelnut Pouncetail", "Raspberry Whiskernibble", "Muffin Nutbounce", "Peach Snickerscuffle",
  "Ginger Fluffpounce", "Biscuit Nuzzlesnack", "Cherry Scurryfluff", "Bluebell Wigglewhisk",
  "Apricot Snugglesnack", "Plum Gigglesnout", "Sunny Snickersnuggle", "Banana Whiskerbounce",
  "Papaya Nutwhisk", "Coconut Snickersnuggle", "Mango Wiggletail", "Berry Gigglesnout",
  "Kiwi Pouncesnout", "Melon Nuzzlenose", "Tangerine Wigglepaws", "Plum Scurrywhisk",
  "Pumpkin Fluffertail", "Grape Nibblesnack", "Blackberry Pouncepaws", "Raspberry Nutwhisk",
  "Peach Fluffertail", "Kiwi Snickersnack", "Mango Giggletail", "Tangerine Fluffernose",
  "Grape Whiskertail", "Blackberry Nutbounce", "Blueberry Scurrypaws", "Cranberry Wiggletail",
  "Mulberry Pouncewhisk", "Orange Nibbler", "Pomegranate Nutsnack", "Strawberry Wigglewhisk",
  "Lemon Nuzzlenose", "Lime Snickerscuffle", "Watermelon Snickersnuggle", "Apple Nutbounce",
  "Cherry Whiskertail", "Grapes Fluffertail", "Peach Wigglepaws", 
  "Sunny", "Chestnut", "Maple", "Peanut", "Hazel", "Walnut", "Oreo", "Marshmallow", "Snickers",  
  "Buttercup","Pumpkin", "Pinecone", "Cocoa", "Blueberry", "Raspberry", "Sparky", "Furry", "Chestnut",
  "Pippin", "Acorn", "Pistachio", "Hazelnut", "Cinnamon", "Mocha", "Pecan", "Acorn", "Biscuit",
  "Almond", "S'mores", "Wiggles", "Twix", "Bubbles", "Doodle", "Wombat", "Fluffernutter",
  "Muffin","Fuzzball", "Cupcake", "Wonton", "Snuggles", "Pippin", "Moonbeam", "Cheesecake", "Jigsaw",
  "Cashew", "Jumper", "Marmalade", "Nimbus", "Toffee", "Poppy", "Squirt", "Munchkin", "Tidbit", 
  "Jingle", "Piston", "Fandango", "Jolt", "Nectar", "Sprinkle", "Nudge", "Tickles", "Hiccup",
  "Sassafras", "Zippy", "Daffodil", "Raisin", "Frizzle", "Caramel", "Stardust", "Quicksilver", "Nishi",
  "Jeff", "Selina", "Ghislaine", "Carolyn", "Camille", "Sanika", "Z", "Will"
)
```



### Export clean data
```{r}
save_path = "clean_squirrel_2018.csv"

write_csv(df_clean_squirrel, save_path, na="")
```
